---
title: "Dummbell"
author: "Richard G. Gardiner"
date: "2/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I will be using the gapminder dataset which is basic economic and social data about a number of countries over multiple year.

## Loading Packages

You will want to load the following packages.

1. tidyverse: has reshaping packages and ggplot2 which is the main source
2. gapminder: a package that includes the gapminder dataset
3. ggalt: another package that enhances the capabilities of ggplot2.  This is where you actually get the geom_dumbbell function
```{r, message = FALSE}
library(tidyverse)
library(gapminder)
library(ggalt)
```

# Loading and Manipulating Data

First lets look at the gapminder dataset:
```{r}
head(gapminder)
```

You can see that each row represents a country-year point with a total of 6 columns.  

Let's say we want a dumbbell plot where we want to look at countries in Europe and the GDP per capita (`gdpPercap`) in 1952 and 1957.  To se that we use the following code:

```{r}
gapminder2 <- gapminder %>% # creates a new object using the gapminder dataset
  filter(year %in% c(1952, 1957)) %>% # filters out all rows that are not in 1952 and 1957
  filter(continent == "Europe") %>% # only want observations in Europe
  select(country, year, gdpPercap) %>% # getting rid of every column that isn't included here
  spread(key = year, value = gdpPercap) # this is where we make each row now a country instead of country-year

head(gapminder2)
```

Now we have data that is now useful for dumbbell plots.  For a dumbbell plot, you need 3 columns.  One categorical variable (state, country, person, etc).  In our example that is `country`.  Then you have two data points that you want to be the end of the dumbbell.  It appears that it is normally a time variable.  In this instance our first end will be `1952` and our second is `1957`.  The code below uses this new dataset, `gapminder2`, to create a dumbbell plot.

# Graphing

First I will make a very simple dumbbell graph which only needs the dataset we are using `gapminder2`, along with the specific variables we have: y = country (this is the categorical variable), x = 1952 (the starting point), xend = 1957 (the ending point).
```{r}
ggplot(data = gapminder2, # calls ggplot and tells it where the data is
       aes(y = country, x = `1952`, xend = `1957`)) + 
  geom_dumbbell()
```

The plot above uses all basic settings in geom_dumbell.  

The one below shows a lot more preferences.  You can really mess around with each setting.  What I generally do is just delete the bottom line and see what happens, then repeat for each line/setting. 

```{r}
ggplot(data = gapminder2, # calls ggplot and tells it where the data is
       aes(y = country, x = `1952`, xend = `1957`)) + # the asthetics: y = CATEGORICAL VARIABLE, x = START_POINT, xend = END_POINT
  
  # EVERYTHING NOW IS JUST SETTINGS AND PREFERENCES
  geom_dumbbell(size=3, color="red",  
                colour_x = "blue", colour_xend = "green",
                dot_guide=FALSE, dot_guide_size=0.25) +
  labs(x=NULL, y=NULL, title="ggplot2 geom_dumbbell with dot guide") +
  theme_minimal() +
  theme(panel.grid.major.x=element_line(size=0.05)) +
  theme(panel.grid.major.y=element_blank())
```

